The role of animal cognition in shaping resource distribution

### 2023-06-16 Text/Article_ModelSeedDispersal.Rmd
Set fixed seed for repro + add moving rules (Rcpp) + Extract results (R)

---
title: "Eco-evolutionary feedback loops between animal cognition, movement and resource distribution"
# The role of animal cognition in shaping resource distribution via zoochory
# date: "`r format(Sys.time(), '%B %d, %Y')`"

#I follow: https://stackoverflow.com/questions/52918716/authors-and-affiliations-in-the-yaml-of-rmarkdown for author display
author: 
  - name: Benjamin Robira
    email: benjamin.robira@normalesup.org
    institute: FEM
    correspondence: true
institute:
  - FEM: Animal Ecology Unit, Research and Innovation Centre, Fondazione Edmund Mach, San Michele all' Adige, TN, Italy

@@

#Updating bibliography
citeR(
bibliographyArticle = "../Text/libraryMdf.bib",
bibliographyOutput = "../Text/bibliographyarticlepackage.bib",
Rcpp,
cccd,
gwfa,
ape,
igraph,
adehabitatLT
)

@@

@@

**Abstract |** Cognition can shape the movement of animals such as for frugivores. As frugivores can be involved in the life cycle of plants, for example in seed dispersal (i.e. zoochory), their cognition could also shape resource distribution patterns. As resource distribution patterns are known to affect the benefits of cognition, this could create eco-evolutionary feedback loops between animal cognition and the spatio-temporal patterns of resources. Using an agent-based model simulating the foraging behaviour of a frugivorous seed disperser endowed with spatio-temporal knowledge of resource distribution, we investigated whether resource spatio-temporal patterns could be influenced by the level of cognition involved in foraging. This level of cognition represented how well resource location and phenology were predicted by the agent.

We showed that ... sentence on the consequence of cognitive foraging on resource distribution  
This stems from the conjunction of two forces: competition for space between trees and a seed-dispersal forager moving from tree to tree based on spatio-temporal memory.  
The engineering of the resource landscape was eventually beneficial/detrimental to the forager...  
Altogether, this study highlights...

@@

\faTags \hspace{0.01in} **Keywords:** Agent-based model - Cognition - Foraging - Frugivory - Movement - Phenology - Seed dispersal

@@

[TO BE UPDATED]

\initial{\textcolor{black}{I}}ntroduce with:

In the theatre of the rainforest, an evolutionary play takes place in an
unchanging setting, or so current hypotheses on the evolution of foragers’ cognition
assume (Rosati, 2017). If the protagonists, the foragers, are likely to have their
cognition shaped by the distribution of resources (Grove, 2013; Robira et al., 2021),
the characteristics of the resources themselves could be the result of forager-induced
selection (Chapman et al., 2013; John et al., 2016; Lomáscolo & Schaefer, 2010). For
example, fruit-eating primates tend to forage on trees that produce the highest quality
fruit (e.g. the sweetest fruit, Ban et al., 2014; Ban et al., 2016) and in the greatest
quantity (Flörchinger et al., 2010). Therefore, the evolutionary history of resources
and associated foragers, as in the angiosperms/primates dyad, may be joint (Fuzessy
et al., 2022; Gómez & Verdú, 2012; Lim et al., 2021; Sussman, 1991).

Lomáscolo, S. B., & Schaefer, H. M. (2010). Signal convergence in fruits: A result of selection
by frugivores? Journal of Evolutionary Biology, 23 (3), 614–624. https://doi.org/10.
1111/j.1420-9101.2010.01931.x

Flörchinger, M., Braun, J., Böhning-Gaese, K., & Schaefer, H. M. (2010). Fruit size, crop
mass, and plant height explain differential fruit choice of primates

Fuzessy, L., Silveira, F. A., Culot, L., Jordano, P., & Verdú, M. (2022). Phylogenetic congruence between Neotropical primates and plants is driven by frugivory. Ecology Letters, 25 (2), 320–329. https://doi.org/10.1111/ele.13918

Gómez, J. M., & Verdú, M. (2012). Mutualism with plants drives primate diversification.
Systematic Biology, 61 (4), 567–577. https://doi.org/10.1093/sysbio/syr127

Sussman, R. W. (1991). Primate origins and the evolution of angiosperms. American Journal
of Primatology, 23 (4), 209–223. https://doi.org/10.1002/ajp.1350230402

Lim, J. Y., Wasserman, M. D., Veen, J., Després-Einspenner, M. L., & Kissling, W. D.
(2021). Ecological and evolutionary significance of primates’ most consumed plant
families. Proceedings of the Royal Society B: Biological Sciences, 288 (1953). https:
//doi.org/10.1098/rspb.2021.0737

§1) The effect of resource distribution on foragers' cognition
-> Cognition enhances foraging movement  
-> Cognition is promoted by specific spatial distribution.  
-> Fix setting (only short term dynamism - because fix during an individual lifetime), but in reality, interaction is two ways

§2) possible consequence of forager on resource distribution
  -> It can affect the spatial distribution:  
  The dispersal way is affecting the distribution.  
  Pooping areas preference -> spatial aggregation  
  The use of memory is associated to the emergence of route, as suggested from theoretical (Louise) to empirical evidence.  
  Route distribution: elephants' route are following gradient. What comes first, following or gradient born from route?  
  -> It can affect the temporal distribution:

Note biblio:
seed dispersal and personality   
https://www.pnas.org/doi/full/10.1073/pnas.2113870119  
https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/1365-2435.13583  
seed dispersal and animal behaviour  
https://link.springer.com/article/10.1007/s00442-005-0178-1  
seed dispersla animal cognition  
https://link.springer.com/article/10.1007/s11258-016-0652-3  
https://onlinelibrary.wiley.com/doi/full/10.1111/j.1469-185X.2012.00250.x  
https://onlinelibrary.wiley.com/doi/full/10.1111/brv.12377  
https://repositories.lib.utexas.edu/bitstream/handle/2152/68127/ADDIS-DISSERTATION-2017.pdf?sequence=1

A la mobilité du frugivore s'oppose la sessilité des ressources.

|   The two-way interaction between frugivore and fruit plants therefore [questionne, necessite une meilleure comprehension] of their joint history [REF]. 
Primates, for instance, have boomed in the late

-> the possible joint history, ref to fuzessy
-> question ISS

In the case of retroaction between frugivore movement, cognition and resource distribution, an extreme scenario is illustrated in Figure \ref{fig:predictions}. Initially, assuming a parcimonious homogeneous distribution, long-term spatio-temporal memory is only marginally advantageous [REF]. As it is nonetheless, cognition should be selected. Through zoochory, deliberate targeting of some resource patches over others should initiate a change in resource distribution (e.g. increasing patchiness), eventually making cognition more and more advantageous. Over time, by repeatedly targeting the same trees, the routes used by the foragers may begin to materialise due to the seeds scattered along the way. However, in doing so, foragers who can only passively follow these routes may also start to become very efficient: they should not even have to bear the cost of high cognitive abilities, putting them at an advantage over foragers bearing the burden of cognition [REF]. While the pattern of resource distribution can be maintained, given that the routes are equally used by all foragers (yet because of two different proximal mechanisms: long-term memory vs. taxis), cognition could even start to be counter-selected. However, in the absence of concrete evidence, such a scenario remains a fable. In this study, we propose to provide a basis for discussing its realism, based on computer simulations that mimic the foraging behaviour of an agent endowed with spatio-temporal knowledge of resource spatial distribution and phenology and the rool of seed-disperser.

```{r predictions, echo = FALSE, message = FALSE, include = TRUE, fig.align = 'center', fig.cap='Theoretical expectations on eco-evolutionary feedback loops between animal movement cognition and resource distribution', out.width = '0.95\\linewidth'}
knitr::include_graphics("../FIG/summaryPredictions.pdf")

@@

The main objective of the model was to simulate the foraging behaviour of a single agent in an environment dynamical both on the short term, through an interplay between resource growth and depletion, and on the long term, due to seed dispersal. For simplicity, we will refer to the agent as the forager, and to resource patches as trees, but this model should find larger resonance. The model was implemented in the *Rcpp* language in *R* software [`r extractRversion()`, @Rsoftware]  based on the *Rcpp* package [@Rcpp1; @Rcpp2; @Rcpp3].

@@

Trees could reproduce and disperse via fruit eaten by the agent. Once the agent ate on ripe fruit, seeds could be dispersed once up to `r timeDelayForDispersal` tu after ingestion, with a probability of `r DispersalProbability/timeDelayForDispersal` tu$^{-1}$, independent of the quantity of fruit ingested, for simplicity. Thus, all fruit eaten did not mandatorily led to dispersal. Dispersal was eventually successful only if some space was left available for the new tree. This means that no tree had to be at a distance lower than `r round(0.80 * sqrt(mapSize * mapSize / (numberTrees * pi)), digits = 2)` su. This number was obtained such as the exclusive "lifespace" by trees (a circular area of radius the aforementioned distance) was equal to 80% of the map area. If it happened that there was available space for the seedling to grow, then the seedling location was sampled in the available space along the linear path linking the previous target. Otherwise, the seedling could not grow and died.

@@

The forager was endowed with a gradient of knowledge abilties: from no memory of the locations of trees (naive forager) and the timing of their period of productivity to perfect omniscience (i.e. all tree locations are known) and prescience (i.e. all periods of productivity are known). In case a tree was part of the forager's memory, the forager knew both the spatial (location) and temporal (fruiting timing) attributes. Thus, it could not know only one or the other. In case of partial knowledge, known trees were randomly selected at start.

<!-- trees growing from successful seedlings were memorised, as well as trees that were visited while fruiting. In this case, they replaced known trees that were visited the longest time ago. To maintain the spatio-temporal knowledge rate constant, each time a new tree location and associate fruiting dates were memorised, those of the tree visited the longest time ago were forgotten. -->

@@

The forager was initially randomly located within the map. It is fully aware of trees present within sensory range fixed to `r round(perceptualRange, digit=2)` su. This actually corresponds to the average nearest-neighbour distance which equates $1/(2\sqrt{\rho})$, with $\rho$ the density of trees (here `r numberTrees/mapSize/mapSize`) at start of the simulation where trees are homogeneously distributed. It is also aware of trees part of its spatio-temporal long-term memory. To decide where to head next, the forager could either rely on its private knowledge (sensory or memory) or not. If all trees within its private knowledge did not yield any fruit (i.e., all having less than 0.001 fu), the forager was moving randomly. Random movements were performed by randomly choosing a direction (uniform distribution of the heading angle), while the step length was sampled within an exponential distribution of mean $\lambda_{step\,length}$ = `r exponentialRate`. An exponential distribution of step lengths is indeed classically observed in nature [REF]. Otherwise, the agent targeted the `r whatRule` tree which maximised the ratio between the predicted food quantity at the tree (which accounts for depletion history), and the distance to the tree. Nonetheless, the forager avoided to target trees recently visited within the previous `r noReturnTime` tu to avoid targeting recently depleted trees [@robirainreview].

@@

|   In case all trees were empty (no food available in the environment, considered if sum of food available was less than 1 fu), the forager entered a torpor period for `r torporTime` tu, and so on until food was available again. This was done to avoid biasing measure of foraging efficiency in case the forager could deplete entirely the environment while being unable to disperse to another area, as expected in nature if this happens.

@@

To assess whether different levels of cognition resulted in different organisation of the spatial distribution of trees at the end of the simulation we focused on three metrics describing the tree spatio-temporal patterns: the *patchiness*, the *alignment* and the *spatial autocorrelation of fruiting timing*.

@@

$P=\frac{N_{quadrat}\sum_{i}{n_{i}(n_{i}-1)}}{(\sum_{i}{n_{i}})^2}$

@@

@@

Routes are (curvi-)linear features. To assess the tendency of trees to form linear pattern, we quantified their alignment. For a given tree, alignment *a* was defined as one minus the minimum of the sine of the angles of the triangle formed by the tree of interest and two closest trees. This varies from perfect alignment (value of 1), to no alignment (i.e., the minimum angle would be at worse 60° hence an alignment of ca. 0.134). Therefore, we use a corrected alignment $a_c$ varying from 0 (no alignment) to 1 (perfect alignment) such as

@@

We used the opposite of the skewness of alignment values of all trees to describe the whole distribution. Positive values indicate the existence of linear motifs while negative values indicate no obvious linear patterns. In the simulated conditions, random distributions (homogeneous or heterogeneous) should highlight moderate positive values [(see Appendix Material)][Using patchiness and alignment to characterise spatial point patterns].

@@

Spatial autocorrelation of fruiting timing (i.e. the tendency of fruiting date of trees to more closely match that of neighbours than of distant trees) was calculated as the Geary Index, \textit{C} [REF] applied to the start of fruiting dates. Since the date of start of fruiting corresponds to a circular variable, we followed [https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5870747/] to adapt existing functions of the *ape* package [@ape] to circular variables. Geary's \textit{C} was therefore defined as

@@

$C = (N-1) \sum_{i}\sum_{j} w_{ij}d(t_i,t_j)^2 \over 2 \sum_{i,j}w_{ij}\sum_i d(x_i,\bar{t})^2$

@@

For each simulation, we verified whether a routine behaviour emerged and paralleled the variations in tree distribution. We characterised routine holding on the notion of conditional entropy following [REF Riotte-Lambert] using the series of tree id targeted during five supplementary season periods in which dispersal could not occur. We computed the routine possibly up to only a first order conditional entropy, as computing routine indices is computationally demanding and a first order is expecting to be the most informative to describe a perfect traplining behaviour [REF Riotte-Lambert]. The routine index varies from 0 (no routine) to 1 (perfect traplining).

### Does resource spatio-temporal distribution vary with cognition level?

To investigate whether different levels of cognition eventually led to different distribution patterns, we considered five forager types differing by the proportion of trees for which the location and phenology was known (from 0 to 100%, at a step of `r 100/length(spatialKnowledge)`, henceforth *spatio-temporal knowledge rate*). Each forager was first let foraging individually for `r cycleLimitNumber` seasonal periods. We repeated the simulations for a given condition (i.e. knowledge rate) `r numberRepetitions` times.

@@

To investigate whether the engineered distribution of trees could in turn affect the foraging benefices purported by cognition, we considered the final tree distributions of simulations computed to answer question 1). For each of these tree distributions, in three different simulation runs, we let forage a forager endowed with no, intermediate (i.e. half of trees known), or full (all trees known) spatio-temporal knowledge. In these simulations, no seed-dispersal could occur. We measured the foraging efficiency after the forager foraged for `r cycleLimitNumber/5` full seasons and compared them across environments and agent abilities. Foraging efficiency was computed as the cumulated quantity of eaten food, over the cumulated distance travelled. We verified that at the end of the simulation that foraging efficiency stabilised, which meant that we reached a foraging equilibrium. We considered that foraging efficiency stabilised when the foraging efficiency sampled after each moving bout during the fourth and fifth quintile of simulation times did not statistically differed of more than `r percentToAssumeDifference`% (reference is the minimum efficiency in the fourth or the fifth quintile of simulation times). Distributions of foraging efficiency measures were compared with Student tests ("t.test" function). We repeated the simulations for a given conditions (i.e. knowledge rate) `r numberRepetitions` times.

### Route formation as a product of two forces: tree competition for space and agents' movement step rule

@@

|   To assess the effect of competition for space between trees, we compared the final tree spatio-temporal patterns when a perfectly knowledgeable agent (all tree locations and phenology known) was let foraging in the same setting as in question 1), with the only difference that the exclusive lifespace between trees could vary be of 5 (small lifespace), 45 (intermediate lifespace) or 85% (large lifespace) of the map surface, so that the radius of this exclusive circular lifespace was at best slightly superior to the perception range (`r round(0.85 * sqrt(mapSize * mapSize / (numberTrees * pi)), digits = 2)` su vs `r round(perceptualRange, digit=2)` su).

|   To assess the effect of the movement tactic of the forager on tree distribution patterns, we compared the final tree spatio-temporal patterns when an omniscient agent was let foraging in the same setting as in question 1) and while the stringency in skipping trees encountered *en route* varied. It could either stop at all trees perceived on the way to the initial target (mimicking a "monitoring" behaviour, as in all other simulations; small skipping stringency), stop only at trees currently fruiting (i.e., with at least 0.001 fu, mimicking an opportunistic feeding only; intermediate skipping stringency) or move directly to the target (large skipping stringency).

@@

[TO BE UPDATED]

```{r figMap, fig.width=7, fig.height=7, include=TRUE, warning = FALSE, fig.cap=""}

# -> add one figure for distribution: 
# - init, after naive agent ran, intermediate or prescient
# - init and three moving rules
# - init and three spacing

@@

```{r figEngineering, fig.width=7, fig.height=7, include=TRUE, warning = FALSE, fig.cap=""}

>>>>

---
title: "The role of animal cognition in shaping resource distribution"
# "Animal cognition as a cause and consequence of resource distribution heterogeneity"
# "Eco-evolutionary feedback loops between animal cognition, movement and resource distribution"
# The role of animal cognition in shaping resource distribution via zoochory
# date: "`r format(Sys.time(), '%B %d, %Y')`"

#I follow: https://stackoverflow.com/questions/52918716/authors-and-affiliations-in-the-yaml-of-rmarkdown for author display
author: 
  - name: Benjamin Robira
    email: benjamin.robira@normalesup.org
    institute: FEM
    correspondence: true
institute:
  - FEM: Animal Ecology Unit, Research and Innovation Centre, Fondazione Edmund Mach, San Michele all'Adige, TN, Italy

@@

#Updating bibliography
citeR(
bibliographyArticle = "../Text/bibliography.bib",
bibliographyOutput = "../Text/bibliographyarticlepackage.bib",
Rcpp,
cccd,
gwfa,
ape,
igraph,
adehabitatLT
)

@@

<!-- When an animal engages in food searching, its cognition can be stimulated, as it may need to remember where and when to find food.  -->

@@

**Abstract |** In the wildlife theater, an evolutionary play takes place in an unchanging setting, or so current hypotheses on the evolution of spatiotemporal cognition of plant-eater animals assume. If these animals are likely to have their spatiotemporal cognition shaped by the distribution of resources, their movement may shape the way plant resources are distributed by dispersing their seeds too. This could create eco-evolutionary feedback loops between animal spatiotemporal cognition and the distribution patterns of plant resources. Using an agent-based model simulating the foraging behaviour of a seed disperser endowed with spatiotemporal knowledge of resource distribution, we investigated whether resource spatiotemporal patterns could be influenced by the level of cognition involved in foraging. This level of cognition represented how well resource location and phenology were predicted by the agent. We showed that seed dispersers could shape the long-term resource distribution with the formation of materialised resource routes. This stemmed from the conjunction of two forces: competition for space between trees and a seed-dispersal forager moving from tree to tree based on spatiotemporal memory. However, the engineering of the resource landscape did not substantially affect the selection gradient that, in turn, could apply to cognition. Altogether, the results emphasise that cognition is a cause and a consequence of resource heterogeneity.

<!-- Cognition can shape the movement of animals such as for frugivores. As frugivores can be involved in the life cycle of plants, for example in seed dispersal (i.e. zoochory), their cognition could also shape resource distribution patterns. As resource distribution patterns are known to affect the benefits of cognition, this could create eco-evolutionary feedback loops between animal cognition and the spatiotemporal patterns of resources.  -->

<!-- Using an agent-based model simulating the foraging behaviour of a seed disperser endowed with spatiotemporal knowledge of resource distribution, we investigated whether resource spatiotemporal patterns could be influenced by the level of cognition involved in foraging. This level of cognition represented how well resource location and phenology were predicted by the agent. We showed that seed dispersers could shape the long-term resource distribution with the formation of materialised resource routes. This stemmed from the conjunction of two forces: competition for space between trees and a seed-dispersal forager moving from tree to tree based on spatiotemporal memory. However, the engineering of the resource landscape did not affect the selection gradient that, in turn, could apply to cognition.  -->

@@

\faTags \hspace{0.01in} **Keywords:** Agent-based model - Feedback loop - Foraging - Frugivory - Memory - Movement - Phenology - Seed dispersal

@@

\initial{\textcolor{black}{P}}erhaps because animal mobility contrasts with plant sessility, hypotheses on the evolution of plant foragers' spatiotemporal cognition generally assume that moving foragers evolve in an unchanging environment over the long term [@rosati2017foraging]. However, while foragers' cognition is likely to be shaped by the distribution of resources [@milton1981distribution; @grove2013evolution; @robiratemporalmemory; @bracis2015memory; @boyer2010modelling], the characteristics of the resources themselves could be the result of forager-induced selection [@valenta2020dispersal; @lomascolo2010signal]. In the tropical rainforests, for example, the cognition of frugivorous primates has probably been shaped by the challenge of the ephemerality and dispersal of their food [@zuberbuhler2010foraging; @trapanese2019review], supporting complex [@teichroeb2019navigation; @trapanese2019whatwhere; @janson2016capuchins; @janson1992measuring; @janmaat2013mistake; @robirarandomwalk] and flexible [@trapanese2022flexibility; @janmaat2006primates] movement heuristics. In turn, primates remember and forage for the most rewarding trees [@ban2016fattyfruit; @ban2014highreward; @florchinger2010fruit]. The latter are thus more likely to have their seeds dispersed, driving plant and landscape evolution [@chapman2013primates].

|   The way seeds are dispersed contributes to the emerging pattern of resource distribution [@vittoz2007seed]. Most of the world's plants [@tiffney1995angiosperm] and especially rainforest trees [main and supplementary material: @gottsberger1983dispersal; @bagchi2011spatial] are dispersed by animals (zoochory). Compared to other modes of dispersal, such as wind (anemochory) or gravity (barychory), zoochory induces a more patchy distribution of adult plants [@seidler2006seed]. This is because animals may tend to revisit memorised sites regularly, and to some extent exclusively [@john2016memoryseed; @addis2017causes]. Furthermore, the imperfection and limited size of spatial memory [@fagan2013spatial] is expected to lead to different movement patterns, and therefore potentially different patterns of resource distribution [@cortes2013integrating]. By being affected by cognition and by affecting resource dispersal, animal movement could thus lead to eco-evolutionary feedback loops between animal cognition and resource dispersal [@john2016cognitionmeetdispersal]. In other words, cognition could be both a cause and a consequence of resource heterogeneity.

|   To illustrate the possible retroactions between cognition and resource distribution, let's consider an extreme speculative scenario in which seed-dispersing frugivores are freely foraging (Figure \@ref(fig:predictions)). At start, assuming a parcimonious homogeneous distribution, long-term spatiotemporal memory is only marginally advantageous [@grove2013evolution]. As it is nonetheless, cognition should be selected for. Through zoochory, deliberate targeting of some resource patches over others may induce a change in resource distribution over time (e.g. increasing patchiness), eventually making cognition more and more advantageous. Gradually, by repeatedly targeting the same trees, the routes used by the foragers may begin to materialise through the seeds scattered along the way. However, in doing so, foragers who can only passively follow these routes and have survived may also begin to become very efficient: they should not even have to bear the cost of high cognitive ability, putting them at an advantage over foragers who bear the burden of cognition [@raichle2006brain]. While the pattern of resource distribution can be maintained because the routes are used equally by all foragers [but through two different mechanisms: long-term memory vs. taxis, @benhamou1995comparative; @benhamou1992distinguishing], cognition could even start to be counter-selected.

|   While some facets of this scenario may be rooted in empirical evidence [e.g. with forest elephants, @campos2011megagardeners; @blake2004fruit; @blake2009forest], most of this scenario remains speculative. The aim of this study is thus to put such a scenario to the test. Using an agent-centred model simulating the movement of seed-dispersing frugivores with different cognitive abilities, we sought to investigate (a) whether the spatial arrangement of resource varies with the cognitive level and movement behaviour of the simulated agents, and (b) whether this in turn affects the selection pressure on the agents' cognition.

```{r predictions, echo = FALSE, message = FALSE, include = TRUE, fig.align = 'center', fig.cap='Theoretical expectations on eco-evolutionary feedback loops between animal spatiotemporal cognition and resource distribution', out.width = '0.95\\linewidth'}
knitr::include_graphics("../FIG/summaryPredictions.pdf")

@@

The main objective of the model was to simulate the foraging behaviour of a single agent in an environment dynamical both on the short term, through an interplay between resource growth and depletion, and on the long term, due to seed dispersal. For simplicity, we will refer to the agent as the forager, and to resource patches as fruit trees, but this model should find larger resonance. The model was implemented in the *Rcpp* language in *R* software [`r extractRversion()`, @Rsoftware]  based on the *Rcpp* package [@Rcpp1; @Rcpp2; @Rcpp3].

@@

Trees could reproduce and disperse via fruit eaten by the agent. Once the agent ate on ripe fruit, seeds could be dispersed once up to `r timeDelayForDispersal` tu after ingestion, with a probability of `r DispersalProbability/timeDelayForDispersal` tu$^{-1}$, independent of the quantity of fruit ingested, for simplicity. Thus, all fruit eaten did not mandatorily led to dispersal. Dispersal was eventually successful only if some space was left available for the new tree. This means that no tree had to be at a distance lower than `r round(0.45 * sqrt(mapSize * mapSize / (numberTrees * pi)), digits = 2)` su. This number was obtained such as the exclusive "lifespace" by trees (a circular area of radius the aforementioned distance) was equal to 45% of the map area (but other percentages were tested, see [Route formation as a product of two forces][Route formation as a product of two forces: tree competition for space and agent movement step rule]). If it happened that there was available space for the seedling to grow, then the seedling location was sampled in the available space along the linear path linking the previous target. Otherwise, the seedling could not grow and died.

@@

The forager was endowed with a gradient of knowledge abilities: from no memory of the locations of trees (naive forager) and the timing of their period of productivity to perfect omniscience (i.e. all tree locations are known) and prescience (i.e. all periods of productivity are known). In case a tree was part of the forager's memory, the forager knew both the spatial (location) and temporal (fruiting timing) attributes. Thus, it could not know only one or the other. In case of partial knowledge, known trees were randomly selected at start.

<!-- trees growing from successful seedlings were memorised, as well as trees that were visited while fruiting. In this case, they replaced known trees that were visited the longest time ago. To maintain the spatiotemporal knowledge rate constant, each time a new tree location and associate fruiting dates were memorised, those of the tree visited the longest time ago were forgotten. -->

@@

The forager was initially randomly located within the map. It is fully aware of trees present within sensory range fixed to `r round(perceptualRange, digit=2)` su. This actually corresponds to the average nearest-neighbour distance which equates $\frac{1}{2\sqrt{\rho}}$, with $\rho$ the density of trees (here `r numberTrees/mapSize/mapSize`) at start of the simulation where trees are homogeneously distributed. It is also aware of trees part of its spatiotemporal long-term memory. To decide where to head next, the forager could either rely on its private knowledge (sensory or memory) or not. If all trees within its private knowledge did not yield any fruit (i.e., all having less than 0.001 fu), the forager was moving randomly. Random movements were performed by randomly choosing a direction (uniform distribution of the heading angle), while the step length was sampled within an exponential distribution of mean $\lambda_{step\,length}$ = `r exponentialRate`. An exponential distribution of step lengths is indeed classically observed in nature [REF]. Otherwise, the agent targeted the `r whatRule` tree which maximised the ratio between the predicted food quantity at the tree (which accounts for depletion history), and the distance to the tree. Nonetheless, the forager avoided to target trees recently visited within the previous `r noReturnTime` tu to avoid targeting recently depleted trees [@robiratemporalmemory].

@@

|   In case all trees were empty (no food available in the environment, considered if the sum of food available was less than 1 fu), the forager entered a torpor period for `r torporTime` tu, and so on until food was available again. This was done to avoid biasing measure of foraging efficiency in case the forager could deplete entirely the environment while being unable to disperse to another area, as expected in nature if this happens.

@@

To assess whether different levels of cognition resulted in different organisation of the spatial distribution of trees at the end of the simulation we focused on three metrics describing the tree spatiotemporal patterns: the *patchiness*, the *alignment* and the *spatial autocorrelation of fruiting timing*.

@@

$P=N_{quadrat}\frac{\sum_{i}{n_{i}(n_{i}-1)}}{(\sum_{i}{n_{i}})^2}$

@@

|   As the forager is set up to forage in a square map with reflective boundaries, seed dispersal may lead to the shrinkage of resource range towards the centre of the map (where recursions are inevitably the higher), an artefact not of interest there. This would contribute to inflate measures of patchiness. We therefore calculated a measure of shrinkage *s* as one minus the area covered by the 95% of the kernel distribution based on all tree locations divided by the map size. To account for variations in shrinkage in the different simulated scenarii [(see Appendix Material)][Variations in resource range shrinkage intensity], we corrected patchiness such as $P_c = P(1-s)$.

@@

Routes are (curvi-)linear features. To assess the tendency of trees to form linear motifs, we quantified their alignment. For a given tree, alignment *a* was defined as one minus the minimum of the sine of the angles of the triangle formed by the tree of interest and the two closest trees. This varies from perfect alignment (value of 1), to no alignment (i.e., the minimum angle would be at worse 60° hence an alignment of ca. 0.134). Therefore, we used a corrected alignment $a_c$ varying from 0 (no alignment) to 1 (perfect alignment) such as

@@

We defined the general alignement for a given environmental map as the opposite of the skewness of alignment values of all trees in the map. Positive values indicate the existence of linear motifs while negative values indicate no obvious linear motifs. Random distributions (homogeneous or heterogeneous) should show moderate positive values, while a route pattern should have a high positive value [(see Appendix Material)][Using patchiness and alignment to characterise spatial point patterns].

@@

Spatial autocorrelation of fruiting timing (i.e. the tendency of fruiting date of trees to more closely match that of neighbours than of distant trees) was calculated as the Geary Index, \textit{C} [@geary1954contiguity] applied to the start of fruiting dates. Since the date of start of fruiting corresponds to a circular variable, we followed [@schmal2017moran] to adapt existing functions of the *ape* package [@ape] to circular variables. Geary's \textit{C} was therefore defined as

@@

$C = \frac{(N-1)}{2} \frac{\sum_{i}\sum_{j} w_{ij}d(t_i,t_j)^2}{\sum_{i,j}w_{ij}\sum_i d(x_i,\bar{t})^2}$

@@

For each simulation, we verified whether a routine behaviour emerged and paralleled the variations in tree distribution. We characterised routine holding on the notion of conditional entropy following [@riotte2016randomness] using the series of the id of trees targeted during five supplementary seasons in which dispersal could not occur. We computed the routine possibly up to only a first order conditional entropy, as computing routine indices is computationally demanding and a first order is expecting to be the most informative to describe a perfect traplining behaviour [@riotte2016randomness]. The routine index varies from 0 (no routine) to 1 (perfect traplining).

### Does resource spatiotemporal distribution vary with cognition level?

To investigate whether different levels of cognition eventually led to different distribution patterns, we considered five forager types differing by the proportion of trees for which the location and phenology was known (from 0 to 100%, at a step of `r 100/length(spatialKnowledge)`, henceforth *spatiotemporal knowledge rate*). Each forager was first let foraging individually for `r cycleLimitNumber` seasonal periods. We repeated the simulations for a given condition (i.e. knowledge rate) `r numberRepetitions` times.

@@

To investigate whether the engineered distribution of trees could in turn affect the foraging benefices purported by cognition, we considered the final tree distributions of simulations computed to answer [question 1][Does resource spatiotemporal distribution vary with cognition level?]). For each of these tree distributions, in three different simulation runs, we let forage a forager endowed with no, intermediate (i.e. half of trees known), or full (all trees known) spatiotemporal knowledge. In these simulations, no seed-dispersal could occur. We measured the foraging efficiency after the forager foraged for `r cycleLimitNumber/5` full seasons and compared them across environments and agent abilities. Foraging efficiency was computed as the cumulated quantity of eaten food, over the cumulated distance travelled. We verified that at the end of the simulation that foraging efficiency stabilised, which meant that we reached a foraging equilibrium. We considered that foraging efficiency stabilised when the foraging efficiency sampled after each moving bout during the fourth and fifth quintile of simulation times did not statistically differed of more than `r percentToAssumeDifference`% (reference is the minimum efficiency in the fourth or the fifth quintile of simulation times). Distributions of foraging efficiency measures were compared with Student tests ("t.test" function). We repeated the simulations for a given conditions (i.e. knowledge rate) `r numberRepetitions` times.

### Route formation as a product of two forces: tree competition for space and agent movement step rule

@@

|   To assess the effect of competition for space between trees, we compared the final tree spatiotemporal patterns when a perfectly knowledgeable agent (all tree locations and phenology known) was let foraging in the same setting as in question 1), with the only difference that the exclusive lifespace between trees could vary be of 5 (small lifespace), 45 (intermediate lifespace) or 85% (large lifespace) of the map surface, so that the radius of this exclusive circular lifespace was always inferior to the perception range (`r round(0.85 * sqrt(mapSize * mapSize / (numberTrees * pi)), digits = 2)` su vs `r round(perceptualRange, digit=2)` su).

|   To assess the effect of the movement tactic of the forager on tree distribution patterns, we compared the final tree spatiotemporal patterns when an omniscient agent was let foraging in the same setting as in question 1) and while the stringency in skipping trees encountered *en route* varied. It could either stop at all trees perceived on the way to the initial target (mimicking a "monitoring" behaviour, as in all other simulations; small skipping stringency), stop only at trees currently fruiting (i.e., with at least 0.001 fu, mimicking an opportunistic feeding only; intermediate skipping stringency) or move directly to the target (large skipping stringency).

@@

<!-- -> For figure engineering, add visualisation to guide meaning of axis: -->

<!-- Y column -->
<!-- schema for each index: -->
<!-- patchiness = aggregated point -->
<!-- route = as in figure pred -->
<!-- spatial auto corr = coloration of points following fruiting date -->
<!-- routine: network of point + segments: in circle = routine, random = no routine -->

<!-- X column: -->
<!-- def of knowledge rate -> Brain size increase, wise number of trees inside -->
<!-- def moving rule: path following trees according to rule -->
<!-- def Space tree: image trees more or less aggregated -->

```{r meanValues}

load("../Renvironment/mainResults.RData")
load("../Renvironment/outputSensitivityMovingRuleAndSpaceTree.RData")

#Getting the mean values
meanValuesMain <- indicesMain_df %>% 
  filter(time == 36500) %>% 
  group_by(knowledgeRate) %>% 
  summarise_all(mean, na.rm = TRUE)

meanValuesEfficiency <- indicesEfficiency_df %>% 
  group_by(initCondition, type) %>% 
  summarise_all(mean, na.rm = TRUE)

meanValuesSpacing <- indicesSpacing_df %>% 
  group_by(valueSpacing) %>% 
  summarise_all(mean, na.rm = TRUE)

meanValuesMovingRule <- indicesMovingRule_df %>% 
  group_by(movingRule) %>% 
  summarise_all(mean, na.rm = TRUE)

#Doing the comparisons of interest
patchinessToNull <- round(meanValuesMain$patchiness/meanValuesMain$patchiness[1], digits = 3)
alignmentToNull <- round(meanValuesMain$alignment/meanValuesMain$alignment[1], digits = 3)

comparisonsEfficiency <- meanValuesEfficiency %>% 
  group_by(initCondition) %>% 
  summarise(
    OmniToIntermediate = round(efficiency[type == "Omniscient"]/efficiency[type == "Intermediate"], digits = 3),
    OmniToNull = round(efficiency[type == "Omniscient"]/efficiency[type == "Null"], digits = 3)
  )

@@

```{r figEngineering, fig.width=18.5, fig.height=18.5, include=TRUE, warning = FALSE, fig.cap="Seed disperser agents can shape resource distribution | The plots show the results of 200 simulations when a seed-dispersing agent with varying level of cognition, moving choice rules or space competition intensities between trees (left to right) was let foraging for a long time. The resulting distribution pattern was assessed by its patchiness (the tendency of trees to be aggregated into patches; larger values indicate highly aggregated trees), its alignement (the tendency of trees to form linear geometries; larger values indicate the presence of linear geometries) and spatial autocorelation in fruiting date (the tendency of trees to have fruiting dates closer in time to their spatial neighbours; values close to 0 indicate negative spatial autocorrelation, values close to 1 indicate no spatial autocorrelation, values above 1 indicate positive spatial autocorrelation). The emerging movement behaviour of the forager was described in terms of routine behaviour (the tendency of the forager to target visit trees in a predictible way; values close to 0 indicate no predictibly movement, values close to 1 indicate highly predictible movements)."}
